Incidence of confirmed COVID-19 cases in care home staff and residents: Cohort study using electronic records in a large UK care home provider (March-June 2020)

### 2020-06-24 covidWarehouse/manuscript/manuscript.Rmd
setting max_date back to 2020-06-07

---
title: "Incidence and period prevalence of confirmed COVID-19 cases in care home staff and residents: Cohort study using electronic records in a large UK care home provider (March-June 2020)"
# author: "Dutey-Magni, PF, Williams, H, Hayward, A, Shallcross, L"
date: "16/06/2020"
bibliography: references.bib
csl: elsevier-vancouver.csl
output:
  html_document:
    df_print: paged
    toc: yes
    code_folding: hide
  word_document:
    toc: yes
    reference_docx: template.docx
link-citations: yes
always_allow_html: yes
---

@@

date_min <- as.Date("2020-03-02")
date_max <- as.Date("2020-06-15")
date_max_tallies = as.Date("2020-06-07")

@@

tline <- dplyr::bind_rows(timelines$timeline)
stopifnot(unique(tline$date_end)>=date_max)
tline <- filter(tline, between(date, date_min, date_max)) %>% 
  mutate(user_type2 = case_when(
    user_type %in% c("General", "Elderly") ~ "General/elderly",
    user_type == "Dementia" ~ "Dementia",
    TRUE ~ NA_character_ ))

@@

homes_country <- select(reference_homes, home_code, postcode) %>% 
  left_join(transmute(reference_geography, postcode, 
                      country, region_england = region), by = "postcode") %>% 
  mutate(region_UK = if_else(country == "England", region_england, country)) %>% 
  select(-postcode)

reference_homes <- left_join(reference_homes, homes_country)

@@

approx_occupancy <- tline %>% 
  group_by(home_code, date) %>% 
  summarise(
    occupancy = sum(tidyr::replace_na(rday, 0)),
    susceptible_symptomatic = sum(susceptible_symptomatic, na.rm = T),
    susceptible_confirmed = sum(susceptible_confirmed, na.rm = T))

approx_occupancy <- merge(
  tibble(date = seq(date_min, date_max, 1)),
  tibble(home_code = home_inclusion),
  all = TRUE) %>% 
  left_join(approx_occupancy)

desc_residents <- tline %>% 
  filter(rday == 1)  %>% 
  group_by(resident_id, home_code, dob, gender, 
           admission_type, home_type, user_type2, status) %>% 
  summarise(resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed)) %>% 
  mutate(ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  left_join(homes_country, by = "home_code")

res_counts <- desc_residents %>% 
  group_by(home_code) %>% 
  summarise(unique_residents = n_distinct(resident_id))

@@

desc_overall <- list()
desc_overall$total_residents <- dplyr::n_distinct(desc_residents$resident_id)
desc_overall$total_staff <- reference_homes %>% filter(home_code %in% unique(residents$home_code)) %>% .$staff_total %>% sum()

@@

@@

desc_overall$total_tests <- nrow(dplyr::filter(incidents, covid_tested == 1))
desc_overall$unique_residents_tested <- dplyr::filter(incidents, covid_tested == 1) %>% 
  dplyr::n_distinct(.$resident_id)
desc_overall$unique_residents_test_result <- dplyr::filter(incidents, !is.na(covid_test_result)) %>% 
  dplyr::n_distinct(.$resident_id)
desc_overall$unique_residents_test_result_positive <- dplyr::filter(incidents, covid_test_result == 1) %>%   dplyr::n_distinct(.$resident_id)

@@

## TABLES
table_nums <- captioner::captioner(prefix = "Table")
tab_national_descriptives <- table_nums(
  name = "tab_national_descriptives",
  caption = "Total homes, beds, and residents within FSHCG and nationwide")
tab_resident_descriptives <- table_nums(
  name = "tab_resident_descriptives",
  caption = paste0("Breakdown of FSHCG residents by care home type, sex, age, region and status on study exit (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
tab_zero_cases <- table_nums(
  name = "tab_zero_cases",
  caption = "Number and proportion of homes without confirmed cases by UK region"
)
tab_CoxPH_confirmed <- table_nums(
  name = "tab_CoxPH_confirmed",
  caption =  paste0("Factors associated with the hazard of become a confirmed case: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(desc_residents$resident_id)),
  ")"))
tab_CoxPH_death <- table_nums(
  name = "tab_CoxPH_death",
  caption =  paste0("Factors associated with all-cause mortality: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(desc_residents$resident_id)),
  ")"))
## FIGURES
fig_nums <- captioner(prefix = "Figure")

fig_KM_CT_residents_symptomatic <- fig_nums(
  name = "fig_KM_CT_residents_symptomatic",
  caption = paste0("Kaplan-Meier estimates of the cumulative probability of residents becoming symptomatic based on care home manager daily counts")
)
fig_KM_residents_confirmed <- fig_nums(
  name = 'fig_KM_residents_confirmed',
  caption = paste0("Kaplan-Meier estimates of the cumulative probability of residents (n=",
  formatbm(dplyr::n_distinct(desc_residents$resident_id)),
  ") becoming confirmed cases based on Datix incidents report"))
fig_KM_CT_confirmed_residents <- fig_nums(
  name = 'fig_KM_CT_confirmed_residents',
  caption = "Kaplan-Meier estimates of the cumulative probability of residents becoming confirmed cases based on care home manager daily counts")
fig_KM_mortality <- fig_nums(
  "fig_KM_mortality",
  caption = paste0("Kaplan-Meier estimates of resident (n=",
  formatbm(dplyr::n_distinct(desc_residents$resident_id)),
  ") survival by SARS-COV-2 case type"))
fig_KM_staff_symptomatic <- fig_nums(
  name = 'fig_KM_staff_symptomatic',
  caption = paste0("Kaplan-Meier estimates of the cumulative probability of care home staff (n=",   formatbm(desc_overall$total_staff),
  ") becoming symptomatic cases based on manager-reported daily counts"))
fig_KM_staff_confirmed <- fig_nums(
  name = 'fig_KM_staff_confirmed',
  caption = paste0("Kaplan-Meier estimates of the cumulative probability of care home staff (n=",   formatbm(desc_overall$total_staff),
  ") becoming confirmed cases based on manager-reported daily counts"))

@@

england_results <- list()
england_results$n_positive = sum(inner_join(new_cases, filter(homes_country, country == "England"))[c("new_confirmed_home", "new_confirmed_hospital")], na.rm = T)
england_results$n_negative = n_distinct(inner_join(residents, filter(homes_country, country == "England"))$resident_id) - england_results$n_positive

@@

**Methods:** Retrospective cohort study using individual-level electronic records from residents (incidents reports) and daily care home counts of infection for both residents and staff. We estimated the incidence and period prevalence of symptoms, confirmed infection, and death in residents and staff, and compare them with findings from a national community survey. We tested a range of risk factors for resident infections in multivariate regression models.

**Findings:** Records were available for `r formatbm(n_distinct(desc_residents$resident_id))` residents present in `r dplyr::n_distinct(residents$home_code)` FSHCG homes between `r format(date_min, "%e %B %Y")` and `r format(date_max, "%e %B %Y")`. The period prevalence of COVID-19 symptoms ranged between `r tbrounding(desc_overall$DAT_p_symptomatic*100)`% (based on incidents reports) and `r tbrounding(desc_overall$CT_p_symptomatic*100)`% (based on manager-reported counts). Across the period, care homes registered a total of `r formatbm(desc_overall$CT_total_confirmed)` (`r tbrounding(desc_overall$CT_p_confirmed*100)`% [`r tbrounding((desc_overall$CT_p_confirmed - desc_overall$CT_p_confirmed_ME)*100)`%; `r tbrounding((desc_overall$CT_p_confirmed + desc_overall$CT_p_confirmed_ME)*100)`%]) infections confirmed with naropharyngeal RNA tests. Incident reports accounted just `r formatbm(desc_overall$unique_residents_test_result_positive)` confirmed infections. In England, `r england_results$n_positive` infections were confirmed of a total `r formatbm(england_results$n_positive + england_results$n_negative)` residents. Comparing with the community, the relative risk of confirmed cases in England was `r tbrounding(RR_england)` [`r tbrounding(exp(log(RR_england) - 1.96 * RR_SE)) `; `r tbrounding(exp(log(RR_england) + 1.96 * RR_SE))`]. The staff-to-bed ratio was the only factor to be independently associated with the rate of confirmed cases: an increase by 10 percentage point in the number of staff relative to the bed total was associated with a 13% reduction [9%, 18%] in the rate of confirmed cases.

**Interpretation** Despite uncertain case ascertainment in care homes, internal surveillance suggests a greater transmission than in the community. Staffing may impact care homes' response to outbreaks and in turn influence rates of confirmed cases. The absence of high-quality case ascertainment hinders local infection prevention initiatives which could be better supported by laboratory data surveillance.

@@

An important source of epidemiological data on the burden of disease in care homes are outbreak investigations.[@McMichael2020] One such investigation estimated that 50% of residents and 20% of staff are infected during an outbreak in London, UK.[@Easter6] Estimates of this nature cannot be generalised to the unknown proportion of care homes that have not reported any outbreaks.

@@

This study obtained data from `r formatbm(dplyr::n_distinct(residents$home_code))` care homes from the Four Seasons Health Care Group (FSHCG), one of the UK's largest independent provider of long-term and respite residential and nursing care. It estimated incidence and period prevalence of COVID-19 events. In addition, it tested and identified individual and organisational risk factors associated with the rate of confirmed infections.

@@

We used individual-level incident reports and daily tallies to generate estimates of disease incidence in residents and period prevalence in residents and staff. Death status as of `r format(date_max, "%e %B %Y")` was used to estimate the case-fatality rate and the relative risks of all-cause death in confirmed cases . The individual data provided an accurate of the duration of exposure in resident-days as well as right-censoring, since individual records indicate when a resident was discharged or died. A Cox proportional hazards model was used to test the effect of potential risk factors, whether individual- (sex, age, care type) or organisation-level (region, care home size, staff-to-bed ratio).

We also investigated whether the number of confirmed infections among staff was associated with subsequent infections amongst residents, independently of time. For this, we regressed.

However, a Datix report may not have been filed for every event. We therefore also estimate the incidence and period prevalence of events using manager counts, using the total resident-days denominator derived from the individual-level data.

The incidence of confirmed cases in England is contrasted with incidence measured from a sample survey of English private households which tested nose/throat swabs from 26,995 individuals between 26 April and 13 June 2020.[@cisbulletin2020; @cisdata2020]

@@

```{r table_resident_descriptives, message = FALSE, results="asis"}
arsenal::tableby(home_type  ~ gender + ageg + region_UK +
                                 user_type2 + admission_type + status, 
                               data = desc_residents) %>% summary()
```

```{r KM_CT_residents_symptomatic, fig.height = 4, fig.width = 7}

@@

For England only, a total of `r england_results$n_positive` confirmed cases occurred out of a total `r formatbm(england_results$n_positive + england_results$n_negative)` residents. In comparison, the survey of English community households[@cisbulletin2020] found a total of `r cis_data$n_positive` confirmed cases out of `r formatbm(cis_data$n_positive + cis_data$n_negative)` sampled individuals. This implies a relative risk of `r tbrounding(RR_england)` [`r tbrounding(exp(log(RR_england) - 1.96 * RR_SE)) `; `r tbrounding(exp(log(RR_england) + 1.96 * RR_SE))`].

```{r km_confirmed, fig.height=4,fig.width=8}

km_confirmed <- tline %>% 
  left_join(select(reference_homes, home_code, postcode, 
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category)) %>% 
  left_join(select(reference_geography, postcode, region_UK)) %>% 
  filter(rday == 1 & susceptible_confirmed == 1) %>% 
  mutate(home_type = factor(home_type, levels = c("Residential", "Nursing"))) %>% 
  group_by(resident_id, dob, gender, region_UK, home_code,
           home_type, user_type2, funding_type, admission_type, 
           staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category) %>% 
  summarise(
    start = as.numeric(min(date) - date_min),
    end = max(date)
  ) %>% 
  left_join(select(timelines, resident_id, covid_first_positive)) %>% 
  mutate(event = if_else(
    is.na(covid_first_positive),
    0L,
    as.integer(covid_first_positive == end)
  )) %>% 
  mutate(
    stop = as.numeric(end +1 - date_min),
    ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)
  )

survfit(Surv(time = start, 
                   time2 = stop,
                   event = event) ~ region_UK, data = km_confirmed) %>% 
  ggsurvplot(., conf.int = TRUE, 
           fun = "event",
           legend = "right",
           ggtheme = theme_bw(),
           # TODO: double check labels at the end
           break.x.by	= 7,
           break.y.by = .02,
           ylim = c(0, .17),
           xlab  = paste0("Days since ", format(date_min, "%d %B %Y")),
           ylab = "Cum probability of confirmed case",
           legend.labs = sort(unique(km_confirmed$region_UK)),
           conf.int.alpha = .1)

@@

```{r KM_CT_residents_infection, fig.height = 4, fig.width = 7}

@@

```{r CPH_confirmed,   results="asis"}
# cph_conf <- coxph(Surv(time = start, time2 = stop, event = event) ~ gender + ageg+home_type+user_type2+region_UK+ staff_to_bed_ratio+bed_tot_category, data = km_confirmed)
# summary(cph_conf)
# cox.zph(cph_conf)
# survfit(Surv(time = start, time2 = stop, event = event)~factor(home_type), data = km_confirmed) %>% ggsurvplot(ylim=c(.85,1), conf.int = T, legend = "right")
explanatory = c("gender", "ageg", "home_type", "user_type2", "region_UK" ,  
                "staff_to_bed_ratio", "bed_tot_category")
dependent = "Surv(time = start, time2 = stop, event = event)"
CPH_confirmed <-  finalfit(km_confirmed, dependent, 
                           explanatory, column = T,
                           add_dependent_label = F)  
knitr::kable(CPH_confirmed,
             row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))
```

@@

```{r risk_death, results="asis"}
# cph_death <- coxph(Surv(date1, date2, death_status) ~
#                       covid_status + gender+ ageg+ home_type +
#                       staff_to_bed_ratio + bed_tot_category +(region_UK)+
#                       cluster(resident_id), data = km_death)
# summary(cph_death)
# cox.zph(cph_death)
# 
explanatory = c("gender", "ageg", "home_type" , 
                "staff_to_bed_ratio", "bed_tot_category", 
                "covid_status", "region_UK", "cluster(resident_id)")
dependent = "Surv(time = date1, time2 = date2, event = death_status)"
CPH_death <-  finalfit(km_death, dependent, explanatory, 
                       column = T, add_dependent_label = F)  
knitr::kable(CPH_death, row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))

@@

```{r KM_Staff_Symptomatic, fig.height = 4, fig.width = 7}

>>>>

---
title: "Incidence of confirmed COVID-19 cases in care home staff and residents: Cohort study using electronic records in a large UK care home provider (March-June 2020)"
# author: "Dutey-Magni, PF, Williams, H, Hayward, A, Shallcross, L"
date: "16/06/2020"
bibliography: references.bib
csl: elsevier-vancouver.csl
output:
  html_document:
    df_print: paged
    toc: yes
    code_folding: hide
  word_document:
    toc: yes
    reference_docx: template.docx
link-citations: yes
always_allow_html: yes
---

@@

date_min <- as.Date("2020-03-02")
date_max <- as.Date("2020-06-07") #as.Date("2020-06-15")
date_max_tallies = as.Date("2020-06-07")

@@

tline <- dplyr::bind_rows(timelines$timeline)
stopifnot(unique(tline$date_end)>=date_max)
tline <- filter(tline, between(date, date_min, date_max)) %>% 
  mutate(home_type = factor(home_type, c("Residential", "Nursing"))) %>% 
  mutate(
    user_type2 = case_when(
      user_type %in% c("General", "Elderly") ~ "General/elderly",
      user_type == "Dementia" ~ "Dementia",
      TRUE ~ NA_character_ 
      ),
    admission_type2 = case_when(
      admission_type %in% c("Continuing Care", "Independent Living") ~ "Continuing care/independent living", 
      TRUE ~ admission_type
    ))

@@

homes_country <- select(reference_homes, home_code, postcode) %>% 
  left_join(transmute(reference_geography, postcode, 
                      country, region_england = region), by = "postcode") %>% 
  mutate(region_UK = if_else(country == "England", region_england, country)) %>% 
  select(-postcode)
reference_homes <- left_join(reference_homes, homes_country)
rm(homes_country)

@@

approx_occupancy <- tline %>% 
  group_by(home_code, date) %>% 
  summarise(
    occupancy = sum(tidyr::replace_na(rday, 0)),
    susceptible_symptomatic = sum(susceptible_symptomatic, na.rm = T),
    susceptible_confirmed = sum(susceptible_confirmed, na.rm = T))
approx_occupancy <- merge(
  tibble(date = seq(date_min, date_max, 1)),
  tibble(home_code = home_inclusion),
  all = TRUE) %>% 
  left_join(approx_occupancy)

residents_episodes <- tline %>% 
  filter(rday==1) %>% 
  group_by(resident_id, home_code, dob, gender, 
           status, home_type, user_type2,
           susceptible_symptomatic, 
           susceptible_confirmed) %>% 
  summarise(entry_date = min(date),
            exit_date = max(date+1),
            resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed)) %>% 
  mutate(date1 = as.numeric( entry_date - date_min),
         date2 = as.numeric(exit_date - date_min),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  mutate(symptomatic = -susceptible_symptomatic+1,
         confirmed = -susceptible_confirmed+1,
         symptomatic_confirmed = (-susceptible_symptomatic+1)*(-susceptible_confirmed+1),
         date2 = if_else(date1 == date2, date2 + 1, date2)) %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Negative",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Negative", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed")),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95))   %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category)) %>% 
  group_by(resident_id) %>% 
  mutate(death_status = if_else(
    date1 == max(date1),
    as.integer(status == "Deceased"),
    0L
  )) %>% 
  ungroup()

residents_overview <- tline %>% 
  filter(rday==1) %>%
  mutate(ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  group_by(resident_id, home_code, ageg, gender, 
           admission_type2, home_type, user_type2, status) %>% 
  summarise(resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed),
            symptomatic = max(-susceptible_symptomatic+1, na.rm = T),
            confirmed = max(-susceptible_confirmed+1, na.rm = T),
            start_date = min(date, na.rm = T),
            stop_date = max(date, na.rm = T),
            start_int = as.numeric(min(date, na.rm = T) - date_min),
            stop_int = as.numeric(max(date, na.rm = T)- date_min)
    ) %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Negative",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Negative", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed"))) %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category)) %>% 
  left_join(distinct(select(incidents, resident_id, starts_with("covid_first"))))

res_counts <- residents_overview %>% 
  group_by(home_code) %>% 
  summarise(unique_residents = n_distinct(resident_id))

@@

desc_overall <- list()
desc_overall$total_residents <- dplyr::n_distinct(residents_overview$resident_id)
desc_overall$total_staff <- reference_homes %>% filter(home_code %in% unique(residents_overview$home_code)) %>% .$staff_total %>% sum()

@@

desc_overall$total_tests <- nrow(dplyr::filter(incidents, covid_tested == 1))
desc_overall$unique_residents_tested <- dplyr::filter(incidents, covid_tested == 1) %>% 
  dplyr::summarise(n = n_distinct(resident_id)) %>% .$n
desc_overall$unique_residents_test_result <- dplyr::filter(incidents, !is.na(covid_test_result)) %>% 
  dplyr::summarise(n = n_distinct(resident_id)) %>% .$n
desc_overall$unique_residents_test_result_positive <- dplyr::filter(incidents, covid_test_result == 1) %>%  
  dplyr::summarise(n = n_distinct(resident_id)) %>% .$n

desc_overall$DAT_p_confirmed <- desc_overall$unique_residents_test_result_positive/desc_overall$total_residents
desc_overall$DAT_p_confirmed_ME <- 1.96 * sqrt(
  desc_overall$DAT_p_confirmed * (1-desc_overall$DAT_p_confirmed) / 
    desc_overall$total_residents
)

desc_overall$DAT_incidence_confirmed <- desc_overall$unique_residents_test_result_positive/sum(residents_overview$susceptible_confirmed_days)
desc_overall$DAT_incidence_confirmed_SE <- sqrt((1-desc_overall$DAT_incidence_confirmed)/desc_overall$unique_residents_test_result_positive)
desc_overall$DAT_incidence_confirmed_95low <- exp(
  log(desc_overall$DAT_incidence_confirmed - 1.96 * desc_overall$DAT_incidence_confirmed_SE)
)
desc_overall$DAT_incidence_confirmed_95up <- exp(
  log(desc_overall$DAT_incidence_confirmed + 1.96 * desc_overall$DAT_incidence_confirmed_SE)
)

@@

@@

## TABLES
table_nums <- captioner::captioner(prefix = "Table")
tab_national_descriptives <- table_nums(
  name = "tab_national_descriptives",
  caption = "Total homes, beds, and residents within FSHCG and nationwide")
tab_resident_descriptives <- table_nums(
  name = "tab_resident_descriptives",
  caption = paste0("Breakdown of FSHCG residents by care home type, sex, age, region and status on study exit (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
tab_zero_cases <- table_nums(
  name = "tab_zero_cases",
  caption = "Number and proportion of homes without confirmed cases by UK region"
)
tab_CoxPH_confirmed <- table_nums(
  name = "tab_CoxPH_confirmed",
  caption =  paste0("Factors associated with the hazard of become a confirmed case: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))
tab_CoxPH_death <- table_nums(
  name = "tab_CoxPH_death",
  caption =  paste0("Factors associated with all-cause mortality: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))
## FIGURES
fig_nums <- captioner(prefix = "Figure")

fig_KM_CT_residents_symptomatic <- fig_nums(
  name = "fig_KM_CT_residents_symptomatic",
  caption = paste0("Kaplan-Meier estimates of the cumulative probability of residents becoming symptomatic based on care home manager daily counts")
)
fig_KM_residents_confirmed <- fig_nums(
  name = 'fig_KM_residents_confirmed',
  caption = paste0("Kaplan-Meier estimates of the cumulative probability of residents (n=",
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ") becoming confirmed cases based on Datix incidents report"))
fig_KM_CT_confirmed_residents <- fig_nums(
  name = 'fig_KM_CT_confirmed_residents',
  caption = "Kaplan-Meier estimates of the cumulative probability of residents becoming confirmed cases based on care home manager daily counts")
fig_KM_mortality <- fig_nums(
  "fig_KM_mortality",
  caption = paste0("Kaplan-Meier estimates of resident (n=",
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ") survival by SARS-COV-2 case type"))
fig_KM_staff_symptomatic <- fig_nums(
  name = 'fig_KM_staff_symptomatic',
  caption = paste0("Kaplan-Meier estimates of the cumulative probability of care home staff (n=",   formatbm(desc_overall$total_staff),
  ") becoming symptomatic cases based on manager-reported daily counts"))
fig_KM_staff_confirmed <- fig_nums(
  name = 'fig_KM_staff_confirmed',
  caption = paste0("Kaplan-Meier estimates of the cumulative probability of care home staff (n=",   formatbm(desc_overall$total_staff),
  ") becoming confirmed cases based on manager-reported daily counts"))

@@

england_results <- list()
england_results$n_positive = sum(inner_join(new_cases, filter(reference_homes, country == "England"))[c("new_confirmed_home", "new_confirmed_hospital")], na.rm = T)
england_results$n_negative = n_distinct(inner_join(residents, filter(reference_homes, country == "England"))$resident_id) - england_results$n_positive

@@

**Methods:** Retrospective cohort study using individual-level electronic records from residents (incidents reports) and daily care home counts of infection for both residents and staff. We estimated the incidence proportion and incidence rate of symptoms, confirmed infection, and death in residents and staff, and compare them with findings from a national community survey. We tested a range of risk factors for resident infections in multivariate regression models.

**Findings:** Records were available for `r formatbm(n_distinct(residents_overview$resident_id))` residents present in `r dplyr::n_distinct(residents$home_code)` FSHCG homes between `r format(date_min, "%e %B %Y")` and `r format(date_max, "%e %B %Y")`. The incidence proportion of COVID-19 symptoms ranged between `r tbrounding(desc_overall$DAT_p_symptomatic*100)`% (based on incidents reports) and `r tbrounding(desc_overall$CT_p_symptomatic*100)`% (based on manager-reported counts). Across the period, care homes registered a total of `r formatbm(desc_overall$CT_total_confirmed)` (`r tbrounding(desc_overall$CT_p_confirmed*100)`% [`r tbrounding((desc_overall$CT_p_confirmed - desc_overall$CT_p_confirmed_ME)*100)`%; `r tbrounding((desc_overall$CT_p_confirmed + desc_overall$CT_p_confirmed_ME)*100)`%]) infections confirmed with nasopharyngeal RNA tests. Incident reports accounted just `r formatbm(desc_overall$unique_residents_test_result_positive)` confirmed infections. In England, `r england_results$n_positive` infections were confirmed of a total `r formatbm(england_results$n_positive + england_results$n_negative)` residents. Comparing with the community, the relative risk of confirmed cases in England was `r tbrounding(RR_england)` [`r tbrounding(exp(log(RR_england) - 1.96 * RR_SE)) `; `r tbrounding(exp(log(RR_england) + 1.96 * RR_SE))`]. The staff-to-bed ratio was the only factor to be independently associated with the rate of confirmed cases: an increase by 10 percentage point in the number of staff relative to the bed total was associated with a 13% reduction [9%, 18%] in the rate of confirmed cases.

**Interpretation** Despite uncertain case ascertainment in care homes, internal surveillance suggests a greater transmission than in the community. Staffing may impact care homes' response to outbreaks and in turn influence rates of confirmed infections The absence of high-quality case ascertainment hinders local infection prevention initiatives which could be better supported by laboratory data surveillance.

@@

An important source of epidemiological data on the burden of disease in care homes are outbreak investigations.[@McMichael2020; @Dora2020] One such investigation estimated that 50% of residents and 20% of staff are infected during an outbreak in London, UK.[@Easter6] Estimates of this nature cannot be generalised to the unknown proportion of care homes that have not reported any outbreaks.

@@

This study obtained data from `r formatbm(dplyr::n_distinct(residents$home_code))` care homes from the Four Seasons Health Care Group (FSHCG), one of the UK's largest independent provider of long-term and respite residential and nursing care. It estimated incidence of SARS-CoV-2 symptoms, diagnoses and all-cause mortality. In addition, it tested and identified individual and organisational risk factors associated with the rate of confirmed infections.

@@

We used individual-level incident reports and daily tallies to generate estimates of infection incidence in residents and staff. The individual data provided an accurate of the duration of exposure in resident-days as well as right-censoring, since individual records indicate when a resident was discharged or died. A Cox proportional hazards model was used to test the effect of potential risk factors, whether individual- (sex, age, care type) or organisation-level (region, care home size, staff-to-bed ratio). Dates of death were used to estimate the case-fatality rate, as well as compare all-cause mortality based on the presence of symptoms and infection confirmation. A time-dependent Cox proportional hazards model analysed the effect of the same factors on all-cause mortality based on the time-dependent infection classification (susceptible; symptomatic not confirmed; asymptomatic confirmed; symptomatic confirmed).

We also investigated whether the number of confirmed infections among staff was associated with subsequent infections amongst residents, independently of time. For this, a Poisson model regressed the rate of confirmed resident cases against home occupancy levels as well as lagged rates of confirmed staff infections at care home level.

A Datix report may not have been filed for every event. We therefore also estimate the incidence of events using manager counts. The resident-days denominator was computed by subtracting past confirmed cases counts from resident-days estimated from the individual-level data.

The incidence proportion of confirmed cases in England is contrasted with incidence measured from a sample survey of English private households which tested nose/throat swabs from 26,995 individuals between 26 April and 13 June 2020.[@cisbulletin2020; @cisdata2020]

@@

```{r table_resident_descriptives, message = FALSE, results="asis"}
arsenal::tableby(home_type  ~ gender + ageg + region_UK +
   user_type2 + admission_type2 + covid_status + status +
   resident_days + susceptible_symptomatic_days + susceptible_confirmed_days, 
   control = arsenal::tableby.control(test = FALSE, numeric.stats = c("Nmiss", "meansd"), digits = 1L),
   data = residents_overview) %>% summary()
```

```{r KM_CT_residents_symptomatic, fig.height = 4, fig.width = 8}

@@

The overall incidence proportion was between `r tbrounding(desc_overall$DAT_p_confirmed*100)`% [`r tbrounding((desc_overall$DAT_p_confirmed - desc_overall$DAT_p_confirmed_ME)*100)`%; `r tbrounding((desc_overall$DAT_p_confirmed + desc_overall$DAT_p_confirmed_ME)*100)`%] according to incidents reports and  `r tbrounding(desc_overall$CT_p_confirmed*100)`% [`r tbrounding((desc_overall$CT_p_confirmed - desc_overall$CT_p_confirmed_ME)*100)`%; `r tbrounding((desc_overall$CT_p_confirmed + desc_overall$CT_p_confirmed_ME)*100)`%]  according to manager counts. For England only, a total of `r england_results$n_positive` confirmed cases occurred out of a total `r formatbm(england_results$n_positive + england_results$n_negative)` residents. In comparison, the survey of English community households[@cisbulletin2020] found a total of `r cis_data$n_positive` confirmed cases out of `r formatbm(cis_data$n_positive + cis_data$n_negative)` sampled individuals. This implies a relative risk of `r tbrounding(RR_england)` [`r tbrounding(exp(log(RR_england) - 1.96 * RR_SE)) `; `r tbrounding(exp(log(RR_england) + 1.96 * RR_SE))`].

```{r km_confirmed, fig.height=4,fig.width=8, warning=F}
survfit(Surv(time = start_int, 
             time2 = stop_int + 1,
             event = confirmed) ~ region_UK, data = residents_overview) %>% 
  ggsurvplot(., conf.int = TRUE, 
           fun = "event",
           legend = "right",
           ggtheme = theme_bw(),
           # TODO: double check labels at the end
           break.x.by	= 7,
           break.y.by = .01,
           ylim = c(0, .08),
           xlab  = paste0("Days since ", format(date_min, "%d %B %Y")),
           ylab = "Cum probability of confirmed case",
           legend.labs = sort(unique(residents_overview$region_UK)),
           conf.int.alpha = .1) 
```

@@

```{r KM_CT_residents_infection, fig.height = 4, fig.width = 8}

@@

```{r CPH_confirmed,   results="asis"}
# cph_conf <- coxph(Surv(time = start_int, time2 = stop_int +1, event = confirmed) ~ gender + ageg+home_type+user_type2+region_UK+ staff_to_bed_ratio+bed_tot_category, data = residents_overview)
# summary(cph_conf)
# cox.zph(cph_conf)
# survfit(Surv(time = start_int, time2 = stop_int +1, event = confirmed)~factor(gender), data = residents_overview) %>% ggsurvplot(ylim=c(.85,1), conf.int = T, legend = "right")
explanatory = c("gender", "ageg", "home_type", "user_type2", "region_UK" ,  
                "staff_to_bed_ratio", "bed_tot_category")
dependent = "Surv(time = start_int, time2 = stop_int +1, event = confirmed)"
CPH_confirmed <-  finalfit(residents_overview,
                           dependent, 
                           explanatory, column = T,
                           add_dependent_label = F)  
knitr::kable(CPH_confirmed,
             row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))
```

@@

```{r risk_death, results="asis"}
cph_death2 <- coxph(Surv(date1, date2, death_status) ~
                      gender+ ageg +
                      staff_to_bed_ratio + bed_tot_category +(region_UK)+ 
                      home_type*covid_status+ cluster(resident_id), data = residents_episodes)
# summary(cph_death)
# cox.zph(cph_death)
# survfit(Surv(time = date1, time2 = date2, event = death_status)~ gender, data = residents_episodes)
# 
explanatory = c("gender", "ageg", "user_type2", "home_type",
                "staff_to_bed_ratio", "bed_tot_category", 
                "covid_status", "region_UK", "cluster(resident_id)")
explanatory_interaction = c("gender", "ageg", "user_type2",
                "staff_to_bed_ratio", "bed_tot_category", 
                "covid_status:home_type", "region_UK", "cluster(resident_id)")
dependent = "Surv(time = date1, time2 = date2, event = death_status)"
CPH_death <-  finalfit(residents_episodes, dependent, explanatory, 
                       column = T, add_dependent_label = F)  
knitr::kable(CPH_death, row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))

@@

```{r KM_Staff_Symptomatic, fig.height = 4, fig.width = 8}
